---
title: 'Assignment 4 - EDS 241: Environmental Policy Evaluation'
author: "Juliet Cohen"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output:
  pdf_document:
    toc: false
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(dplyr)
library(estimatr)
library(knitr)
library(stargazer)
library(tidyverse)
library(AER)
library(lmtest)
library(car)
library(lfe)
```

Read in data:

```{r}
data <- read.csv("/Users/juliet/Documents/MEDS/EDS_241_Env_Policy_Eval/Assignments/Assignment_4/EDS241_Assignment_4/EU_sardines.csv")
```


This question will ask you to estimate the price elasticity of demand for fresh sardines across **56 ports** located in **4 European countries** with **monthly** data from **2013 to 2019**. The data are contained in the file EU_sardines.csv, which is available on Gauchospace. Each row in the data file is a combination of port location (where the fish is landed and sold) in a given year and month. You can ignore the fact that the sample is not balanced (the number of monthly observations varies across ports).

For the assignment, you will need the following variables: **year, month, country, port (port where sardines are landed and sold), price_euro_kg (price per kg in €), and volume_sold_kg (quantity of sardines sold in kg).** In the questions below, I use `log()` to denote the natural logarithm.

# (a) Estimate a bivariate regression of log(volume_sold_kg) on log(price euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.

```{r}
data_log <- data %>% 
  mutate(log_volume_sold_kg = log(volume_sold_kg),
         log_price_euro_kg = log(price_euro_kg))

model_price_vol <- lm(log_volume_sold_kg ~ log_price_euro_kg, data = data_log)
summary(model_price_vol)

linearHypothesis(model_price_vol, c("log_price_euro_kg=-1"), white.adjust = "hc2")
```

\noindent - **The price elasticity of demand for sardines is `r model_price_vol$coefficients[[2]]`, meaning that if the log price per kg of sardines (in €) increases by 1 percent, the log volume sold (in kg) decreases by `r model_price_vol$coefficients[[2]]` percentage points. We can reject the null hypothesis that the price elasticity of demand is equal to -1 because the p-value for this coefficient is smaller than 0.05, so this is a statistically significant result, and the linear hypothesis test shows that it is significant.**

# (b) Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic.

```{r}
# first stage regression
model_price_wind <- lm(log_price_euro_kg ~ wind_m_s, data = data_log)
summary(model_price_wind)

# generate F-statistic
linearHypothesis(model_price_wind, c("wind_m_s=0"), white.adjust = "hc2")
```

\noindent - **The estimated coefficient for wind speed is `r model_price_wind$coefficients[[2]]`, which represents the percent point increase in log price per kg (in €) for every 1 unit increase in wind speed (1 m/s). Yes, this coefficient has the expected sign, because I presume that more windy conditions make fishing more difficult, which would result in the cost of sardines to increase. The F-statistic is 144.65, which is greater than 10, indicating that this is a relevant and non-weak instrument.**

# (c) Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines?

```{r}
tsls_vol_price_wind <- ivreg(log_volume_sold_kg ~ log_price_euro_kg | wind_m_s, data = data_log)
summary(tsls_vol_price_wind)
```

\noindent - **The estimated price elasticity of demand for sardines is `r tsls_vol_price_wind$coefficients[[2]]`, which represents the percent point decrease in log volume of sardines sold (in kg) for every 1 percent increase in log price per kg (in €), using wind as an instrument.**

# (d) Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint: you can use the command “as.factor(country) + as.factor(year) + as.factor(month)” to the ivreg function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.

```{r}
fe_model <- lm(log_price_euro_kg ~ as.factor(country) + as.factor(year) 
               + as.factor(month) + wind_m_s, data = data_log)
summary(fe_model)

linearHypothesis(fe_model, c("wind_m_s=0"), white.adjust = "hc2")

# same regression with fixed effects for year, month, and country
fe_tsls_vol_price_wind <- ivreg(log_volume_sold_kg ~ log_price_euro_kg +
                               + as.factor(country) + as.factor(year) 
                               + as.factor(month) | as.factor(country) 
                               + as.factor(year) + as.factor(month) 
                               + wind_m_s, data = data_log)
# ivreg() is running 2 first stage least regressions, first with the coefficients on the right side 
# of | and then with the coefficients on the left
summary(fe_tsls_vol_price_wind)
```


\noindent - **The estimated price elasticity of demand is -1.250 with fixed effects for month, year, and country. The F-statistic is 77.658, which is greater than 10, so this instrument is non-weak and relevant.**







